Two applications of logic programming to Coq

### 2020-11-05 papers/types2020/sec3.tex
Explain prolog tactic

\todo{explain Ev = Term as an Elpi thing}

@@

\begin{lstlisting}
Elpi Tactic dprolog.
Elpi Accumulate lp:{{
  solve [str ''height'', int N] [goal _  Ev G _] _ :-
    coq.say "Goal:" {coq.term->string G},
    check (qheight N) (go G Term),
    Ev = Term,
    coq.say "Proof:" {coq.term->string Ev}.
...
(* Other clauses for different fpc omitted *)
}}.
\end{lstlisting}
The following example show how we can use the above tactic to do
FPC-driven logic programming in Coq and return a Coq proof-term:
\begin{lstlisting}
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : is_nat x -> insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, y <= x -> insert x ys rs -> insert x (y :: ys) (x :: rs).
Lemma i1:  exists R, insert 2 [0;1] R.
eexists.
elpi  dprolog height 10.
Qed.
Print i1.
ex_intro (fun R : list nat => insert 2 [0; 1] R) [0; 1; 2]
  (i_c 2 0 [1] [1; 2] (i_c 2 1 [] [2] (i_n 2)) : exists R : list nat, insert 2 [0; 1] R
\end{lstlisting}

>>>>

@@

\begin{lstlisting}
Elpi Tactic dprolog.
Elpi Accumulate lp:{{
  solve [str ''height'', int N] [goal _  Ev G _] _ :-
    coq.say "Goal:" {coq.term->string G},
    check (qheight N) (go G Term),
    Ev = Term,
    coq.say "Proof:" {coq.term->string Ev}.
...
(* Other clauses for different fpc omitted *)
}}.
\end{lstlisting}

The glue code between Coq-Elpi and the implementation of our calculus is
straightforward: the goal is given to us as a quadruple of a context, an
evar, a type and a list of extra information. In additon, the certificate is
supplied by the user when calling the tactic (it simply is an integer in this
case). We just need to call \lsti{check} on the goal's type, together with
the certificate, in order to obtain a reconstructed proof term. We avoid
calling the reconstruction directly on the evar because Coq-Elpi ensures that
evars manipulated by \lP are well-typed at all times, and we cannot guarantee
that since we work with partially reconstructed term; getting around this is
simply a matter of assigning the reconstructed term to the evar right
afterwards.
The following example shows how we can use the above tactic to do
FPC-driven logic programming in Coq and return a Coq proof-term:
\begin{lstlisting}
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : is_nat x -> insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, y <= x -> insert x ys rs -> insert x (y :: ys) (x :: rs).
Lemma i1:  exists R, insert 2 [0;1] R.
eexists.
elpi  dprolog height 10.
Qed.
Print i1.
ex_intro (fun R : list nat => insert 2 [0; 1] R) [0; 1; 2]
  (i_c 2 0 [1] [1; 2] (i_c 2 1 [] [2] (i_n 2)) : exists R : list nat, insert 2 [0; 1] R
\end{lstlisting}
